---
title: "Cleveland Housing Study-Population"
date: "03/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label = "allow")

options(scipen=3)
options(knitr.table.format = "latex")


library(tidycensus)
library(tidyverse)
library(plotly)
library(censusapi)
library(censusGeography)
library(dplyr)
library(DT)


# my census API key: "75789d63464567dcba5b8edb7500fe6148ff3440"

# Add key to .Renviron
Sys.setenv(CENSUS_KEY="75789d63464567dcba5b8edb7500fe6148ff3440")

# Reload .Renviron
readRenviron("~/.Renviron")
# Check to see that the expected key is output in your R console

Sys.getenv("CENSUS_KEY")

myvar = "B01001_001"

cnty_citypop <- get_acs(
  geography = "county subdivision", 
  variables = c(totalpop=myvar), 
  state = "OH", 
  county = "Cuyahoga",
  year = 2019,
  output = "wide"
)%>%
  mutate(NAME = str_remove(NAME, ", Cuyahoga County, Ohio"))

cnty_citypop
cnty_citypop = cnty_citypop[cnty_citypop$GEOID > "3903500000",]
cnty_citypop


cnty_citypop10 <- get_acs(
  geography = "county subdivision", 
  variables = c(totalpop10=myvar), 
  state = "OH", 
  county = "Cuyahoga",
  year = 2010,
  output = "wide"
)

cnty_citypop10 =select(cnty_citypop10, 1, 3:4)
cnty_citypop10

cnty_citypop15 <- get_acs(
  geography = "county subdivision", 
  variables = c(totalpop15=myvar), 
  state = "OH", 
  county = "Cuyahoga",
  year = 2015,
  output = "wide"
)

cnty_citypop15 =select(cnty_citypop15, 1, 3:4)
cnty_citypop15


#---------------- how to compare two ACH --------------------------------------------------
mypop <- merge(cnty_citypop15, cnty_citypop, by = "GEOID") # NA's match, so 6 rows
mypop

mypop$popchange <-round((mypop$totalpopE-mypop$totalpop15E)/mypop$totalpopE * 100, 2)
mypop


mypop10 <- merge(cnty_citypop10, cnty_citypop, by = "GEOID") # NA's match, so 6 rows
mypop10

mypop10$popchange <-round((mypop10$totalpopE-mypop10$totalpop10E)/mypop$totalpopE * 100, 2)
mypop10



myplot <- ggplot(mypop, aes(x = popchange, y = reorder(NAME, popchange))) + 
  geom_col() + 
  labs(title = "Population Percent Change from 2015 to 2019", 
       subtitle = "Cities in Cuyahoga", 
       x = "Percentage Change (%)", 
       y = "") + 
  theme_minimal(base_size = 9)+
  geom_vline(xintercept=-1.38, color="orange", size=1) 



myplot10 <- ggplot(mypop10, aes(x = popchange, y = reorder(NAME, popchange))) + 
  geom_col() + 
  labs(title = "Population Percent Change from 2010 to 2019", 
       subtitle = "Cities in Cuyahoga", 
       x = "Percentage Change (%)", 
       y = "") + 
  theme_minimal(base_size = 9)+
  geom_vline(xintercept=-6.21, color="orange", size=1) 

mytable <- merge(mypop, mypop10, by = "GEOID") 

#-------------------- select fields ------------------------------------------------
#dplyr


mytable <- mytable %>%
  select(NAME.x, totalpop10E, totalpop15E, totalpopE.y, popchange.y, popchange.x)

names(mytable)[1] <- 'City'
names(mytable)[2] <- 'TotalPop2010'
names(mytable)[3] <- 'TotalPop2015'
names(mytable)[4] <- 'TotalPop2019'
names(mytable)[5] <- '%Change10-19'
names(mytable)[6] <- '%Change15-19'


```

The city of Cleveland has a long-term population shrinking along with the rust belt cities in northeast Ohio. Cleveland lost 6.21 % of the total population between 2010 and 2019, ranked 6th city in Cuyahoga County. The population loss was 1.38% between 2015 and 2019. It was the 27th rank among 58 cities in Cuyahoga County. Past five years, Cleveland city has experienced less population loss than in 2010 to 2015.

```{r echo=FALSE, results="asis"}

 mytable%>%
  datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('excel', 'pdf', 'print'),
                           lengthMenu = list(c(25,50,-1),
                                             c(25,50,"All"))),
            rownames = FALSE,
            caption ='population chang from 2010, 2015 to 2019')
  

```



```{r pressure, echo=FALSE}
myplot10

```



```{r presure, echo=FALSE}

myplot


```

Source: Census ACH 5 Year Estimate 2010, 2015, and 2019
Published: CD MIS & GIS Office 
